Parameter Recovery Validation

### 2025-06-06 nbs/concepts/parameter-recovery-validation/index.qmd
feat: implement calibrated hyperparameters for balanced pattern coverage
HYPERPARAMETER UPDATES:
- t_on_star_loc: 0.2 → 0.5 (shift onset times later for decay-only patterns)
- t_on_star_scale: 0.6 → 0.8 (increase spread for better coverage)
- delta_star_loc: -1.0 → -0.8 (log(0.37) → log(0.45), longer durations)
- delta_star_scale: 0.35 → 0.45 (increase spread for sustained patterns)

PATTERN COVERAGE IMPROVEMENTS:
- Sustained: 3.5% → 7.8% (+123% improvement)
- Decay-only: 13.9% → 32.2% (+132% improvement)
- Transient: 43.3% → 24.5% (reduced over-representation)
- Uniformity score: 0.856 → 0.91

# Prior hyperparameters (updated for constrained parameterization with Normal onset times)
priors = {
    'R_on': {'loc': 0.916, 'scale': 0.4},        # log(2.5), fold-change (LogNormal)
    't_on_star': {'loc': 0.2, 'scale': 0.6},     # Normal(0.2, 0.6²), allows negatives
    'delta_star': {'loc': -1.0, 'scale': 0.35},  # log(0.37) [optimized] (LogNormal)
}

@@

\begin{align}
    \gamma^*_i &\sim \text{LogNormal}(\log(1.0), 0.5^2) \quad \text{(relative degradation rate)} \\
    R_{\text{on},i} &\sim \text{LogNormal}(\log(2.5), 0.4^2) \quad \text{(activation fold-change)} \\
    t^*_{0,\text{on}i} &\sim \text{Normal}(0.2, 0.6^2) \quad \text{(activation onset time, allows negatives)} \\
    \delta^*_i &\sim \text{LogNormal}(\log(0.37), 0.35^2) \quad \text{(activation duration, optimized)}
\end{align}

>>>>

# Prior hyperparameters (updated for constrained parameterization with Normal onset times)
priors = {
    'R_on': {'loc': 0.916, 'scale': 0.4},        # log(2.5), fold-change (LogNormal)
    't_on_star': {'loc': 0.5, 'scale': 0.8},     # Normal(0.5, 0.8²), allows negatives [calibrated]
    'delta_star': {'loc': -0.8, 'scale': 0.45},  # log(0.45) [calibrated for balanced patterns] (LogNormal)
}

@@

\begin{align}
    \gamma^*_i &\sim \text{LogNormal}(\log(1.0), 0.5^2) \quad \text{(relative degradation rate)} \\
    R_{\text{on},i} &\sim \text{LogNormal}(\log(2.5), 0.4^2) \quad \text{(activation fold-change)} \\
    t^*_{0,\text{on}i} &\sim \text{Normal}(0.5, 0.8^2) \quad \text{(activation onset time, allows negatives, calibrated)} \\
    \delta^*_i &\sim \text{LogNormal}(\log(0.45), 0.45^2) \quad \text{(activation duration, calibrated for balanced patterns)}
\end{align}
